### 2019-07-25 misc/proofs/coprime-probability.tex
Less than 2x+1

\proof To prove this limit, we will investigate the difference
$$q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 = \sum_{k\geq 1}\mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2.$$
Because $\lfloor n/k \rfloor = 0$ for $k>n$, the first summation on the right-hand side need only range up to $k=n$, and we can reorganise the summations as follows:
$$\eqalignno{
    \sum_{k\geq 1}\mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n \mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k=1}^n \mu(k)\left({n\over k}\right)^2 - \sum_{k>n}\mu(k)\left({n\over k}\right)^2\cr
    q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n \mu(k) \left(\left\lfloor {n\over k} \right\rfloor ^2 - \left({n \over k}\right)^2\right) - \sum_{k>n}\mu(k)\left({n\over k}\right)^2 & ({\oldstyle 1})
}$$
We will want to put asymptotic bounds on these rearranged sums. First, note that for all positive real $x$,
$$x^2 - \lfloor x\rfloor^2 = O(x).\oldno 2$$
For $x^2 - \lfloor x\rfloor ^2$ can be rewritten $(x + \lfloor x\rfloor)(x-\lfloor x\rfloor)$, which is less than $2x$. Next, we use standard formulas to analyse the asymptotic behaviour of the series $\sum_{k>n}(n/k)^2$:
$$\eqalign{
     \sum_{k>n} \left({n\over k}\right)^2 &= \sum_{k\geq 1} \left({n\over k}\right)^2 - \sum_{k=1}^n \left({n\over k}\right)^2 \cr
    &= {n^2\pi^2\over 6} - n^2\left({\pi^2\over 6} -  {1\over n} + O( n^{-2} )\right) \cr
    &= n - O(1) \cr
    &= O(n)
}\oldno 3$$
Since $\mu(k)\leq 1$, we can discard this factor when applying Eqs.\ ({\oldstyle 2}) and ({\oldstyle 3}) to get the asymptotic behaviour of Eq.\ ({\oldstyle 1}); this yields
$$\eqalign{
    q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n O(n/k) - O(n)\cr
    &= O(nH_n) - O(n),
}$$
where $H_n = \sum_{k=1}^n 1/k$ denotes the $n$-th harmonic number. Because both $O(nH_n)$ and $O(n)$ are $o(n^2)$, we can replace these terms by $o(n^2)$ and divide by $n^2$ to arrive at
$${q_n\over n^2} - \sum_{k\geq 1}{\mu(k)\over k^2} = o(1).$$
This means that for any real $\epsilon > 0$, we can find a number $n_0$ such that for all $n \geq n_0$ the difference $|q_n - \sum_{k\geq 1} \mu(k)/k^2| \leq \epsilon$; the sequence $q_n/n^2$ converges exactly as stated by the lemma.\slug

>>>>

\proof To prove this limit, we will investigate the difference
$$q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 = \sum_{k\geq 1}\mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2.$$
Because $\lfloor n/k \rfloor = 0$ for $k>n$, the first summation on the right-hand side need only range up to $k=n$, and we can reorganise the summations as follows:
$$\eqalignno{
    \sum_{k\geq 1}\mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n \mu(k)\left\lfloor {n\over k}\right\rfloor^2 - \sum_{k=1}^n \mu(k)\left({n\over k}\right)^2 - \sum_{k>n}\mu(k)\left({n\over k}\right)^2\cr
    q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n \mu(k) \left(\left\lfloor {n\over k} \right\rfloor ^2 - \left({n \over k}\right)^2\right) - \sum_{k>n}\mu(k)\left({n\over k}\right)^2 & ({\oldstyle 1})
}$$
We will want to put asymptotic bounds on these rearranged sums. First, note that for all positive real $x$,
$$x^2 - \lfloor x\rfloor^2 = O(x).\oldno 2$$
For $x^2 - \lfloor x\rfloor ^2$ can be rewritten $(x + \lfloor x\rfloor)(x-\lfloor x\rfloor)$, which is less than $2x+1$. Next, we use standard formulas to analyse the asymptotic behaviour of the series $\sum_{k>n}(n/k)^2$:
$$\eqalign{
     \sum_{k>n} \left({n\over k}\right)^2 &= \sum_{k\geq 1} \left({n\over k}\right)^2 - \sum_{k=1}^n \left({n\over k}\right)^2 \cr
    &= {n^2\pi^2\over 6} - n^2\left({\pi^2\over 6} -  {1\over n} + O( n^{-2} )\right) \cr
    &= n - O(1) \cr
    &= O(n)
}\oldno 3$$
Since $\mu(k)\leq 1$, we can discard this factor when applying Eqs.\ ({\oldstyle 2}) and ({\oldstyle 3}) to get the asymptotic behaviour of Eq.\ ({\oldstyle 1}); this yields
$$\eqalign{
    q_n - \sum_{k\geq 1}\mu(k)\left({n\over k}\right)^2 &= \sum_{k=1}^n O(n/k) - O(n)\cr
    &= O(nH_n) - O(n),
}$$
where $H_n = \sum_{k=1}^n 1/k$ denotes the $n$-th harmonic number. Because both $O(nH_n)$ and $O(n)$ are $o(n^2)$, we can replace these terms by $o(n^2)$ and divide by $n^2$ to arrive at
$${q_n\over n^2} - \sum_{k\geq 1}{\mu(k)\over k^2} = o(1).$$
This means that for any real $\epsilon > 0$, we can find a number $n_0$ such that for all $n \geq n_0$ the difference $|q_n - \sum_{k\geq 1} \mu(k)/k^2| \leq \epsilon$; the sequence $q_n/n^2$ converges exactly as stated by the lemma.\slug
